---
title: "Thesis project - "
output: html_notebook
---

## Setup environment 

### Clean environment and set directory
```{r setup}
rm(list=ls())

knitr::opts_knit$set(root.dir = normalizePath("C:\\Users\\beasav\\OneDrive - KI.SE\\Bioinformatics\\Thesis"))

```

### Libraries
```{r}
library("xlsx") #Read/Write excel files
library(ggrepel) #plots
library(clusterProfiler)
library(org.Hs.eg.db)
```

---
# Enrichr_KEGG_2021
## Load data

SARS-CoV-2 UN_48H
SARS-CoV-1 UN48_48H
MERS-CoV UN_24H

```{r}
CoV2 <- read.xlsx("SARS-CoV-2 UN_48H\\Data and figures\\Enrichr_KEGG_2021_sig.xlsx",1)
CoV1 <- read.xlsx("SARS-CoV-1 UN_48H\\Data and figures\\Enrichr_KEGG_2021_sig.xlsx",1)
MERS <- read.xlsx("MERS-CoV UN_24H\\Data and figures\\Enrichr_KEGG_2021_sig.xlsx",1)
```

## Venn diagram
```{r}
library(VennDiagram)

# Chart
venn.diagram(
  x = list(CoV1$Term, MERS$Term, CoV2$Term),
  category.names = c("SARS-CoV-1", "MERS-CoV", "SARS-CoV-2"),
  filename = 'Venn_diagram_KEGG_2021.png',
  output=TRUE,
  
  #Title
  main = "KEGG 2021",
  main.pos = c(0.5 , 1),
  main.fontfamily = "sans",
  main.cex = 2,
  main.just = c(0.5, 0.5),
  
  imagetype="png",
          #height = 480 , 
          #width = 3000 , 
          #resolution = 300,
          #compression = "lzw",
          lwd = 1,
          col=c('#70abb3', "#f37736", "#7bc043"),
          fill = c(alpha('#70abb3',0.3), alpha("#f37736",0.3), alpha("#7bc043",0.3)),
          cex = 1.5,
          fontfamily = "sans",

          cat.cex = 1.5,
          cat.default.pos = "outer",
          cat.pos = c(-170, 170, 0),
          cat.dist = c(0.025, 0.025, 0.02),
          cat.fontfamily = "sans",
          cat.col = c('#70abb3', "#f37736", "#7bc043"),
          rotation = 1
)
```

## Find common pathways
### Vectors with the indexes of the common pathways
```{r}

CoV2_CoV1 = c()
CoV2_MERS = c()
CoV1_MERS = c()
All = c()

for (i in 1:nrow(CoV2)){
  if (CoV2$Term[i] %in% CoV1$Term){
    CoV2_CoV1 = c(CoV2_CoV1, CoV2$Term[i])
  }
  if (CoV2$Term[i] %in% MERS$Term){
    CoV2_MERS = c(CoV2_MERS, CoV2$Term[i])
  }
}

for (i in 1:nrow(CoV1)){
  if (CoV1$Term[i] %in% MERS$Term){
    CoV1_MERS = c(CoV1_MERS, CoV1$Term[i])
  }
}

for (i in 1:length(CoV1_MERS)){
  if (CoV1_MERS[i] %in% CoV2_MERS && CoV1_MERS[i] %in% CoV2_MERS){
    All = c(All, CoV1_MERS[i])
  }
}

rm(i)

```

### Data.frames with info about common pathways from the different viruses

```{r}
CoV2_common <- data.frame(Term = c(), Overlap = c(), Adjusted.P.value = c(), Genes = c())
CoV1_common <- data.frame(Term = c(), Overlap = c(), Adjusted.P.value = c(), Genes = c())
MERS_common <- data.frame(Term = c(), Overlap = c(), Adjusted.P.value = c(), Genes = c())


for (i in match(All, CoV2$Term)){
  C2 <- data.frame(Term = CoV2$Term[i], Overlap = CoV2$Overlap[i], Adjusted.P.value = CoV2$Adjusted.P.value[i], Genes = CoV2$Genes[i])
  CoV2_common <- rbind(CoV2_common, C2)
  
}

for (i in match(All, CoV1$Term)){
  C1 <- data.frame(Term = CoV1$Term[i], Overlap = CoV1$Overlap[i], Adjusted.P.value = CoV1$Adjusted.P.value[i], Genes = CoV1$Genes[i])
  CoV1_common <- rbind(CoV1_common, C1)
  
}
for (i in match(All, MERS$Term)){
  CM <- data.frame(Term = MERS$Term[i], Overlap = MERS$Overlap[i], Adjusted.P.value = MERS$Adjusted.P.value[i], Genes = MERS$Genes[i])
  MERS_common <- rbind(MERS_common, CM)
  
}

#To make it the right order for ggplot

CoV2_common$position <- "SARS-CoV-2"
CoV2$Term <- factor(CoV2$Term, levels = CoV2$Term )

CoV1_common$position <- "SARS-CoV-1"
CoV1$Term <- factor(CoV1$Term, levels = CoV1$Term )

MERS_common$position <- "MERS-CoV"
MERS$Term <- factor(MERS$Term, levels = MERS$Term )


rm(C2)
rm(C1)
rm(CM)
```

## Dot plot

```{r}
plot <- rbind(CoV2_common, CoV1_common, MERS_common)
plot$position <- factor(plot$position, levels = c("SARS-CoV-2", "SARS-CoV-1", "MERS-CoV"))

pdf("Common_pathways_enrichr_KEGG_2021_dot.pdf", width = 6, height = 3)

p<-ggplot(plot, aes(x = position, y = Term)) + 
  geom_point(aes(size = Overlap, color = Adjusted.P.value)) + 
  scale_color_gradientn(colors = c("#668466", "#CC0000")) + 
  theme_classic() + 
  theme(axis.title.y=element_blank(), axis.title.x=element_blank(), axis.ticks.x=element_blank(), legend.key.size = unit(0.3, 'cm'), legend.title = element_text(size = 8),  legend.text = element_text(size = 6)) 

p

dev.off()

rm(p)
rm(plot)

```



# Heatmap, coronavirus disease
## Load KEGG_2021 hsa05171 - Coronavirus disease pathway genes

http://rest.kegg.jp/get/hsa05171

```{r}
genes <- read.xlsx("KEGG_hsa05171_Coronavirus disease_genes.xlsx", 1, header = TRUE)
genes <- genes$Gene_name
```


## Load Abundance data from all viruses

```{r}
CoV2_abundance <- read.xlsx("SARS-CoV-2 UN_48H\\Data and figures\\CoV2_UN_48_filtered.xlsx",1)
CoV1_abundance <- read.xlsx("SARS-CoV-1 UN_48H\\Data and figures\\CoV1_UN_48_filtered.xlsx",1)
MERS_abundance <- read.xlsx("MERS-CoV UN_24H\\Data and figures\\MERS_UN_24_filtered.xlsx",1)

```


## Matrix with logFC values

```{r}

data <- data.frame(CoV2 = rep(NaN, length(genes)), CoV1 = rep(NaN, length(genes)), MERS = rep(NaN, length(genes)))

rownames(data) <- genes

CoV2_genes <- c()
CoV1_genes <- c()
MERS_genes <- c()

#row.names(data) <- genes

for (i in 1:length(genes)){
  
  #Go through CoV2
  for (l in 1:nrow(CoV2_abundance)){
    if (CoV2_abundance$adj.P.Val[l] < 0.05 & genes[i] == CoV2_abundance$Gene_name[l]){
      data$CoV2[i] <- CoV2_abundance$logFC[l]
      
      CoV2_genes <- c(CoV2_genes, genes[i])
      
    }
  }
  
  #Go through CoV1
  for (l in 1:nrow(CoV1_abundance)){
    if (CoV1_abundance$adj.P.Val[l] < 0.05 & genes[i] == CoV1_abundance$Gene_name[l]){
      data$CoV1[i] <- CoV1_abundance$logFC[l]
      
      CoV1_genes <- c(CoV1_genes, genes[i])
    }
  }
  
  #Go through MERS
  for (l in 1:nrow(MERS_abundance)){
    if (MERS_abundance$adj.P.Val[l] < 0.05 & genes[i] == MERS_abundance$Gene_name[l]){
      data$MERS[i] <- MERS_abundance$logFC[l]
      
      MERS_genes <- c(MERS_genes, genes[i])
    }
  }
  
}

#data[is.na(data)] <- 0

rm(i)
rm(l)
```

### Write genes from this pathway in files for STRING

```{r}
write(CoV2_genes, "STRING\\CoV2_genes.txt")
write(CoV1_genes, "STRING\\CoV1_genes.txt")
write(MERS_genes, "STRING\\MERS_genes.txt")
```


### Heatmap_All

```{r}
library("gplots")

pdf("Heatmat_coronavirus disease_all.pdf", width = 6, height = 20)

heatmap.2(as.matrix(data), Rowv = FALSE,
          Colv = FALSE,
          dendrogram = "none",
          tracecol=NA,
          na.color = "#E0E0E0",
          scale = "none",
          col = colorRampPalette(c("#003300", "#325b32","white", "#F2490C", "#670000")),
          
          #Axis labels
          cexRow = 0.1,
          cexCol = 1.1,
          srtCol = 45,
          labCol = c("SARS-CoV-2", "SARS-CoV-1", "MERS-CoV"),
          
          
          # color key + density info
          #key = TRUE,
          #keysize = 5,
          density.info="none",
          #denscol=tracecol,
          #symkey = any(x < 0, na.rm=TRUE) || symbreaks,
          #densadj = 0.25,
          key.title = NA,
          key.xlab = NA, #"logFC",
          #key.ylab = NULL,
          #key.xtickfun = NULL,
          #key.ytickfun = NULL,
          #key.par = list(mar=c(10,10,3,0)),
          
          
          # plot layout
          #lmat = 1,
          lhei = c(0.5, 10),
          lwid = c(3, 5),
          margins=c(6,9)
)

dev.off()

```
### Heatmap_All_nolabel

```{r}
library("gplots")

pdf("Heatmat_coronavirus disease_all_nobels.pdf", width = 6, height = 20)

heatmap.2(as.matrix(data), Rowv = FALSE,
          Colv = FALSE,
          dendrogram = "none",
          tracecol=NA,
          na.color = "#E0E0E0",
          scale = "none",
          col = colorRampPalette(c("#003300", "#325b32","white", "#F2490C", "#670000")),
          
          #Axis labels
          #cexRow = 0.1,
          labRow = NA,
          cexCol = 1.1,
          srtCol = 45,
          labCol = c("SARS-CoV-2", "SARS-CoV-1", "MERS-CoV"),
          
          
          # color key + density info
          #key = TRUE,
          #keysize = 5,
          density.info="none",
          #denscol=tracecol,
          #symkey = any(x < 0, na.rm=TRUE) || symbreaks,
          #densadj = 0.25,
          key.title = NA,
          key.xlab = NA, #"logFC",
          #key.ylab = NULL,
          #key.xtickfun = NULL,
          #key.ytickfun = NULL,
          #key.par = list(mar=c(10,10,3,0)),
          
          
          # plot layout
          #lmat = 1,
          lhei = c(0.5, 10),
          lwid = c(3, 5),
          margins=c(6,9)
)

dev.off()

```
### Heatmap with only differentially regulated in at least one of the viruses
#### Clean NA rows
```{r}

data_clean <- data.frame(CoV2 = c(), CoV1 = c(), MERS = c())

genes_clean <- c()

for (i in 1:nrow(data))
  if (!(is.na(data$CoV2[i]) & is.na(data$CoV1[i]) & is.na(data$MERS[i]))){
    genes_clean <- c(genes_clean, genes[i])
    d <- data.frame(CoV2 = data$CoV2[i], CoV1 = data$CoV1[i], MERS = data$MERS[i])
    data_clean <- rbind(data_clean, d)
  }

rownames(data_clean) <- genes_clean

rm(d)
rm(i)
rm(genes_clean)

```

#### Heatmap clean data_key_nolabel
```{r}
library("gplots")

pdf("Heatmat_coronavirus disease_clean_key_nolabel.pdf", width = 5, height = 8)

heatmap.2(as.matrix(data_clean),
          Rowv = FALSE,
          Colv = FALSE,
          dendrogram = "none",
          tracecol=NA,
          na.color = "#E0E0E0",
          scale = "none",
          col = colorRampPalette(c("#003300", "#325b32","white", "#F2490C", "#670000")),
          
          #Axis labels
          #cexRow = 1,
          labRow = NA,
          cexCol = 1.2,
          srtCol = 45,
          labCol = c("SARS-CoV-2", "SARS-CoV-1", "MERS-CoV"),
          
          
          # color key + density info
          key = TRUE,
          #keysize = 5,
          density.info="none",
          #denscol=tracecol,
          #symkey = any(x < 0, na.rm=TRUE) || symbreaks,
          #densadj = 0.25,
          key.title = NA,
          key.xlab = NA, #"logFC",
          #key.ylab = NULL,
          #key.xtickfun = NULL,
          #key.ytickfun = NULL,
          key.par = list(mar=c(2,0.5,0.5,0.5)), #bottom,left,top,right
          
          
          # plot layout
          #lmat = 1,
          lhei = c(1, 13),
          lwid = c(2, 5),
          margins=c(6,0.5)
)

dev.off()
```
#### Heatmap clean data_nolabel
```{r}
library("gplots")

pdf("Heatmat_coronavirus disease_clean_nolabel.pdf", width = 3, height = 5)

heatmap.2(as.matrix(data_clean),
          Rowv = FALSE,
          Colv = FALSE,
          dendrogram = "none",
          tracecol = NA,
          na.color = "#E0E0E0",
          scale = "none",
          col = colorRampPalette(c("#003300", "#325b32","white", "#F2490C", "#670000")),
          
          #Axis labels
          #cexRow = 1,
          labRow = NA,
          cexCol = 1,
          srtCol = 45,
          labCol = c("SARS-CoV-2", "SARS-CoV-1", "MERS-CoV"),
          
          
          # color key + density info
          key = FALSE,
          #keysize = 5,
          #density.info="none",
          #denscol=tracecol,
          #symkey = any(x < 0, na.rm=TRUE) || symbreaks,
          #densadj = 0.25,
          #key.title = NA,
          #key.xlab = NA, #"logFC",
          #key.ylab = NULL,
          #key.xtickfun = NULL,
          #key.ytickfun = NULL,
          #key.par = list(mar=c(2,0.5,0.5,0.5)), #bottom,left,top,right
          
          
          # plot layout
          #lmat = 1,
          lhei = c(0.1, 13),
          lwid = c(1, 7),
          margins=c(5,0.5)
)

dev.off()
```
---
# Cytoscape

## Get cytoscape tables
```{r}

CoV2_cytoscape <- read.xlsx("STRING\\CoV2_cytoscape_table_filtered.xlsx",1)

CoV1_cytoscape <- read.xlsx("STRING\\CoV1_cytoscape_table_filtered.xlsx",1)

MERS_cytoscape <- read.xlsx("STRING\\MERS_cytoscape_table_filtered.xlsx",1)


```


## Fill cytoscape tables
```{r}
for (i in 1:nrow(CoV2_cytoscape)){
  for (l in 1:nrow(CoV2_abundance)){
    if (CoV2_cytoscape$display.name[i] == CoV2_abundance$Gene_name[l]){
      CoV2_cytoscape$adj.P.Val[i] <- CoV2_abundance$adj.P.Val[l]
      CoV2_cytoscape$logFC[i] <- CoV2_abundance$logFC[l]
    }
  }
}

for (i in 1:nrow(CoV1_cytoscape)){
  for (l in 1:nrow(CoV1_abundance)){
    if (CoV1_cytoscape$display.name[i] == CoV1_abundance$Gene_name[l]){
      CoV1_cytoscape$adj.P.Val[i] <- CoV1_abundance$adj.P.Val[l]
      CoV1_cytoscape$logFC[i] <- CoV1_abundance$logFC[l]
    }
  }
}

for (i in 1:nrow(MERS_cytoscape)){
  for (l in 1:nrow(MERS_abundance)){
    if (MERS_cytoscape$display.name[i] == MERS_abundance$Gene_name[l]){
      MERS_cytoscape$adj.P.Val[i] <- MERS_abundance$adj.P.Val[l]
      MERS_cytoscape$logFC[i] <- MERS_abundance$logFC[l]
    }
  }
}

```

## write in file
```{r}

write.xlsx(CoV2_cytoscape, "STRING\\CoV2_cytoscape_table.xlsx", col.names = TRUE, row.names = FALSE)
write.xlsx(CoV1_cytoscape, "STRING\\CoV1_cytoscape_table.xlsx", col.names = TRUE, row.names = FALSE)
write.xlsx(MERS_cytoscape, "STRING\\MERS_cytoscape_table.xlsx", col.names = TRUE, row.names = FALSE)
```

