---
title: "Thesis project - "
output: html_notebook
---

## Setup environment 

### Clean environment and set directory
```{r setup}
rm(list=ls())

knitr::opts_knit$set(root.dir = normalizePath("C:\\Users\\beasav\\OneDrive - KI.SE\\Bioinformatics\\Thesis"))

```

### Libraries
```{r}
library("xlsx") #Read/Write excel files
library(ggrepel) #plots
library(clusterProfiler)
library(org.Hs.eg.db)
```

---
# Enrichr_KEGG_2021
## Load data

SARS-CoV-2 UN_48H
SARS-CoV-1 UN48_48H
MERS-CoV UN_24H

```{r}
CoV2 <- read.xlsx("SARS-CoV-2 UN_48H\\Data and figures\\Enrichr_KEGG_2021_sig.xlsx",1)
CoV1 <- read.xlsx("SARS-CoV-1 UN_48H\\Data and figures\\Enrichr_KEGG_2021_sig.xlsx",1)
MERS <- read.xlsx("MERS-CoV UN_24H\\Data and figures\\Enrichr_KEGG_2021_sig.xlsx",1)
```

## Venn diagram
```{r}
library(VennDiagram)

# Chart
venn.diagram(
  x = list(CoV1$Term, MERS$Term, CoV2$Term),
  category.names = c("SARS-CoV-1", "MERS-CoV", "SARS-CoV-2"),
  filename = 'Venn_diagram_KEGG_2021.png',
  output=TRUE,
  
  #Title
  main = "KEGG 2021",
  main.pos = c(0.5 , 1),
  main.fontfamily = "sans",
  main.cex = 2,
  main.just = c(0.5, 0.5),
  
  imagetype="png",
          #height = 480 , 
          #width = 3000 , 
          #resolution = 300,
          #compression = "lzw",
          lwd = 1,
          col=c('#70abb3', "#f37736", "#7bc043"),
          fill = c(alpha('#70abb3',0.3), alpha("#f37736",0.3), alpha("#7bc043",0.3)),
          cex = 0.9,
          fontfamily = "sans",

          cat.cex = 1,
          cat.default.pos = "outer",
          cat.pos = c(-170, 170, 0),
          cat.dist = c(0.025, 0.025, 0.02),
          cat.fontfamily = "sans",
          cat.col = c('#70abb3', "#f37736", "#7bc043"),
          rotation = 1
)
```

## Find common pathways
### Vectors with the indexes of the common pathways
```{r}

CoV2_CoV1 = c()
CoV2_MERS = c()
CoV1_MERS = c()
All = c()

for (i in 1:nrow(CoV2)){
  if (CoV2$Term[i] %in% CoV1$Term){
    CoV2_CoV1 = c(CoV2_CoV1, CoV2$Term[i])
  }
  if (CoV2$Term[i] %in% MERS$Term){
    CoV2_MERS = c(CoV2_MERS, CoV2$Term[i])
  }
}

for (i in 1:nrow(CoV1)){
  if (CoV1$Term[i] %in% MERS$Term){
    CoV1_MERS = c(CoV1_MERS, CoV1$Term[i])
  }
}

for (i in 1:length(CoV1_MERS)){
  if (CoV1_MERS[i] %in% CoV2_MERS && CoV1_MERS[i] %in% CoV2_MERS){
    All = c(All, CoV1_MERS[i])
  }
}

rm(i)

```

### Data.frames with info about common pathways from the different viruses

```{r}
CoV2_common <- data.frame(Term = c(), Overlap = c(), Adjusted.P.value = c(), Genes = c())
CoV1_common <- data.frame(Term = c(), Overlap = c(), Adjusted.P.value = c(), Genes = c())
MERS_common <- data.frame(Term = c(), Overlap = c(), Adjusted.P.value = c(), Genes = c())


for (i in match(All, CoV2$Term)){
  C2 <- data.frame(Term = CoV2$Term[i], Overlap = CoV2$Overlap[i], Adjusted.P.value = CoV2$Adjusted.P.value[i], Genes = CoV2$Genes[i])
  CoV2_common <- rbind(CoV2_common, C2)
  
}

for (i in match(All, CoV1$Term)){
  C1 <- data.frame(Term = CoV1$Term[i], Overlap = CoV1$Overlap[i], Adjusted.P.value = CoV1$Adjusted.P.value[i], Genes = CoV1$Genes[i])
  CoV1_common <- rbind(CoV1_common, C1)
  
}
for (i in match(All, MERS$Term)){
  CM <- data.frame(Term = MERS$Term[i], Overlap = MERS$Overlap[i], Adjusted.P.value = MERS$Adjusted.P.value[i], Genes = MERS$Genes[i])
  MERS_common <- rbind(MERS_common, CM)
  
}

#To make it the right order for ggplot

CoV2_common$position <- "SARS-CoV-2"
CoV2$Term <- factor(CoV2$Term, levels = CoV2$Term )

CoV1_common$position <- "SARS-CoV-1"
CoV1$Term <- factor(CoV1$Term, levels = CoV1$Term )

MERS_common$position <- "MERS-CoV"
MERS$Term <- factor(MERS$Term, levels = MERS$Term )


rm(C2)
rm(C1)
rm(CM)
```

## Dot plot

```{r}
plot <- rbind(CoV2_common, CoV1_common, MERS_common)
plot$position <- factor(plot$position, levels = c("SARS-CoV-2", "SARS-CoV-1", "MERS-CoV"))

pdf("Common_pathways_enrichr_KEGG_2021_dot.pdf", width = 7, height = 5)

p<-ggplot(plot, aes(x = position, y = Term)) + geom_point(aes(size = Overlap, color = Adjusted.P.value)) + scale_color_gradientn(colors = c("#668466", "#CC0000")) + theme_classic() + theme(axis.title.x=element_blank(), axis.ticks.x=element_blank()) 

p

dev.off()

rm(p)
rm(plot)

```



# Heatmap, coronavirus disease
## Load KEGG_2021 hsa05171 - Coronavirus disease pathway genes

http://rest.kegg.jp/get/hsa05171

```{r}
genes <- read.delim("KEGG_hsa05171_Coronavirus disease_genes.txt", header = FALSE)
genes <- genes$V1
```


## Load Abundance data from all viruses

```{r}
CoV2_abundance <- read.xlsx("SARS-CoV-2 UN_48H\\Data and figures\\CoV2_UN_48_filtered.xlsx",1)
CoV1_abundance <- read.xlsx("SARS-CoV-1 UN_48H\\Data and figures\\CoV1_UN_48_filtered.xlsx",1)
MERS_abundance <- read.xlsx("MERS-CoV UN_24H\\Data and figures\\MERS_UN_24_filtered.xlsx",1)

```


## Matrix with logFC values

```{r}

data <- data.frame(Genes = genes, CoV2 = rep(NA, length(genes)), CoV1 = rep(NA, length(genes)), MERS = rep(NA, length(genes)))

for (i in 1:length(genes)){
  
  #Go through CoV2
  for (l in 1:nrow(CoV2_abundance)){
    if (CoV2_abundance$adj.P.Val[l] < 0.05 & genes[i] == CoV2_abundance$Gene_name[l]){
      data$CoV2[i] <- CoV2_abundance$logFC[l]
    }
  }
}
```




### Heatmap
```{r}
library(ComplexHeatmap)

pdf("Heatmat_neogi_covid_interferon_2.pdf", width = 10, height = 18)
ht = Heatmap(as.matrix(input[,-c(1, 4, 11:17)]),
             col = col_fun1,
             show_row_names = TRUE,
             column_order = colnames(input)[-c(1, 4, 11:17)],
             show_column_names = FALSE,
             column_split = condition[-3],
             row_split = input$Pathway,row_title_rot = 0,
             column_title_gp = gpar(fill = c(c_t0, c_t24, c_t48, c_t72)),
             border = FALSE,
             column_gap = unit(2, "mm"), 
             right_annotation = rowAnnotation(Mock_H24 = input$LogFC_0_24,
                                              Mock_H48 = input$LogFC_0_48,
                                              H24_H48 = input$LogFC_24_48,
                                              
                                              col = list(
                                                         Mock_H24 = col_fun_2,
                                                         Mock_H48 = col_fun_2,
                                                         H24_H48 = col_fun_2
                                                         ),  show_legend = TRUE, gp = gpar(col = "grey")
             ))
#hm = Heatmap(as.matrix(input[,c(15:17)]), col = col_fun, cluster_rows = FALSE, cluster_columns = FALSE , column_names_rot = 0, column_names_side = "top")
ht_list = ht
draw(ht_list)
dev.off()
```




---