---
title: "Thesis project - SARS-CoV-2 UN_48H"
output: html_notebook
---

## Setup environment 

### Clean environment and set directory
```{r setup}
rm(list=ls())

knitr::opts_knit$set(root.dir = normalizePath("C:\\Users\\beasav\\OneDrive - KI.SE\\Bioinformatics\\Thesis\\MERS-CoV UN_24H\\Data and figures"))
```

### Libraries
```{r}
library("xlsx") #Read/Write excel files
library(ggrepel) #plots
library(clusterProfiler)
library(org.Hs.eg.db)
```


## Load data
MERS-CoV UN48_48
```{r}
data_raw <- read.xlsx("MERS_UNINF_vs_MERS_DGE_results.xlsx",1)

```
### Filter for important info and significance

```{r}

data <- data.frame(Gene_name = rep(NA, nrow(data_raw)), logFC = rep(NA, nrow(data_raw)), adj.P.Val = rep(NA, nrow(data_raw)), Significance = rep(NA, nrow(data_raw)))
#create a Color column

data$Gene_name <- data_raw$Gene_name
data$logFC <- data_raw$logFC
data$adj.P.Val <- data_raw$adj.P.Val

rm(data_raw)

for (i in 1:nrow(data)){
  if(data$logFC[i] > 0.3 & data$adj.P.Val[i] < 0.05){
    data$Significance[i] <- "up_regulated"
  }
  else if(data$logFC[i] < -0.3 & data$adj.P.Val[i] < 0.05){
    data$Significance[i] <- "down_regulated"
  }
  else if(data$adj.P.Val[i] > 0.05){
    data$Significance[i] <- "NS" #Non significant
    #data[i, 3] <- NA #Delete gene name for NS - So that it doesn't appear in the figure?
  }
  else{ #Label remaining significant genes
    data$Significance[i] <- "significant"
    #data[i, 3] <- NA #Delete gene name for significant - So that it doesn't appear in the figure?
  }
}
rm(i)
```

### Save in file
```{r}
write.xlsx(data, "MERS_UN_24_filtered.xlsx", row.names = FALSE)
```


---
## Building Vulcano plot

### Clean labels for nonsignificant and significant genes
```{r}
data_vol <- data

for (i in 1:nrow(data_vol)){
  if(data_vol$Significance[i] == "NS"){
    data_vol$Gene_name[i] <- NA #Delete gene name for NS - So that it doesn't appear in the figure?
  }
  else if(data_vol$Significance[i] == "significant") { #Label remaining significant genes
    data_vol$Gene_name[i] <- NA #Delete gene name for significant - So that it doesn't appear in the figure?
  }
    
}
rm(i)
```

### Volcano plot ggplot

To change the backgorund: Ggplots theme *Check this*
https://www.datanovia.com/en/blog/ggplot-themes-gallery/

Original color pallet: "#003300", "#325b32","#668466","#ffffcc","#670000"
Alternative color pallet: "#006633", "#00994C","#CC6600", "#CC0000", "#990000"

```{r}

pdf("volcano_plot.pdf", height = 5, width = 7)

vol <- ggplot(data = data_vol, aes(x = logFC, y = -log10(adj.P.Val), colour = logFC)) + geom_point(size=1, alpha = 3, na.rm = T) + scale_color_gradientn(colors = c("#003300", "#325b32","#668466","#F2490C","#670000")) + theme(legend.title=element_text(size=10), legend.text=element_text(size=6),
        legend.key.size=unit(0.7,"line"), plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=10), axis.title.x=element_text(size=10),
        axis.text.y=element_text(size=10), axis.text.x=element_text(size=10)) + labs(x="Log2 Fold Change", y="-log10 (adj.P.Val)") + geom_text_repel(aes(label = Gene_name), size=2.5, color = "black") + theme_light() 

vol

dev.off()
```

#### Add horizontal line for p<0.05 in volcano plot and vertival line for logFC <-0.3 and >0.3 in volcano plot

```{r}
pdf("volcano_plot_line.pdf", height = 5, width = 7)

vol <- ggplot(data = data_vol, aes(x = logFC, y = -log10(adj.P.Val), colour = logFC)) + geom_point(size=1, alpha = 3, na.rm = T) + scale_color_gradientn(colors = c("#003300", "#325b32","#668466","#F2490C","#670000")) + theme(legend.title=element_text(size=10), legend.text=element_text(size=6),
        legend.key.size=unit(0.7,"line"), plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=10), axis.title.x=element_text(size=10),
        axis.text.y=element_text(size=10), axis.text.x=element_text(size=10)) + labs(x="Log2 Fold Change", y="-log10 (adj.P.Val)") + geom_text_repel(aes(label = Gene_name), size=2.5, color = "black") + theme_light() + geom_hline(yintercept= - log10(0.05), linetype="dashed") + geom_vline(xintercept = 0.3, linetype="dashed") + geom_vline(xintercept = -0.3, linetype="dashed")

vol

dev.off()

#ggsave("volcano_plot_line.png", vol, height = 5, width = 7)

rm(vol)
rm(data_vol)
```

---
## Enrichr - KEGG 2021 Human
### Filter for significant changes

```{r}
sig_Gene_names <- data$Gene_name[data$Significance != "NS"] #Gene names are stored in raw data; data$Significance indicates significance
```

### File with dif_expressed Gene names for Enrichr
There's one "Uncharacterized protein" - A0A1W2PPE3 https://www.uniprot.org/uniprot/A0A1W2PPE3
Since there is no information about the gene it corresponds to, it is not included for Gene Enrichment Analysis

```{r}
write(sig_Gene_names, "dif_expressed genes.txt")

```

### Import Enrichr results

https://maayanlab.cloud/Enrichr/enrich?dataset=c772337dc1494106be1825a7eae88c48

Library: KEGG 2021 Human

"The p-value is computed from the Fisher exact test which is a proportion test that assumes a binomial distribution and independence for probability of any gene belonging to any set."

*The smaller the p-value, the higher the chance that the genes belong to that pathway, right?*

```{r}
enrichr <- read.delim("Enrichr_KEGG_2021_Human_table.txt")

```

### Filter for only important info
```{r}
enrichr_sig <- data.frame(Term = c(), Overlap = c(), Adjusted.P.value = c(), Genes = c())

for (i in 1:nrow(enrichr)){
  if (enrichr$Adjusted.P.value[i] < 0.05) {
    sig = data.frame (Term = enrichr$Term[i], Overlap = enrichr$Overlap[i], Adjusted.P.value = enrichr$Adjusted.P.value[i], Genes = enrichr$Genes[i])
    enrichr_sig <- rbind(enrichr_sig, sig)
  }
}

enrichr <- enrichr_sig

rm(enrichr_sig)
rm(i)
rm(sig)


```


### Convert overlap into number and round to 2 decimal places

```{r}
for (i in 1:nrow(enrichr)){
  enrichr$Overlap[i] <- round(eval(parse(text = enrichr$Overlap[i])), digits = 2) #2 decimal places
}
rm(i)
#and order from top to bottom overlap
#top <- top[with(top, order(Overlap)),]

```

### Write into file

```{r}
write.xlsx(enrichr, "Enrichr_KEGG_2021_sig.xlsx", sheetName = "Enrichr_sig", col.names= TRUE, row.names = FALSE)
```



### Select the top100 pathways
According to Flora the p-value tells more than the overlap.

enrichr dataset is ordered based on p-value
```{r}
enrichr_top <- data.frame(Term = c(), Overlap = c(), Adjusted.P.value = c())


# do top enrichr but with Interferon and pvalue < 0.05
for ( i in 1:100) {
  top_pathway <- data.frame(Term = enrichr_sig$Term[i], Overlap = enrichr_sig$Overlap[i], Adjusted.P.value = enrichr_sig$Adjusted.P.value[i] )
    
  enrichr_top <- rbind(enrichr_top, top_pathway)
}

rm(top_pathway)

```

#### Building dot plot
```{r}
#order from low to high p-value
enrichr_top$CoV2_UN_48 <- 1
enrichr_top$Term <- factor(enrichr_top$Term, levels = enrichr_top$Term ) #To make it the right order for ggplot
```


```{r}

pdf("Top_Comparison_pathways_enrichr_dot.pdf", width = 18, height = 20)
p<-ggplot(enrichr_top, aes(x = CoV2_UN_48, y = Term)) + geom_point(aes(size = Overlap, color = Adjusted.P.value)) + scale_color_gradientn(colors = c("#668466", "#CC0000")) + theme_classic()

p

dev.off()

rm(p)

```
#### Building top100 bar plot

```{r}
enrichr_top$Term <- factor(enrichr_top$Term, levels = enrichr_top$Term ) #To make it the right order for ggplot

pdf("Comparison_pathways_enrichr_top100_bar.pdf", width = 18, height = 20)
p<-ggplot(data = enrichr_top, aes(x = Term, y = Overlap)) + geom_bar(stat="identity", aes(fill = Adjusted.P.value)) + scale_fill_gradientn(colors = c("#668466", "#CC0000")) + coord_flip() + theme_light() + theme(axis.title.y=element_blank())

p

dev.off()

rm(p)

```
### Select the IFN, ub and proteosome pathways
According to Flora the p-value tells more than the overlap.

enrichr dataset is ordered based on p-value
```{r}
enrichr_IFN <- data.frame(Term = c(), Overlap = c(), Adjusted.P.value = c())


# do top enrichr but with Interferon and pvalue < 0.05
for ( i in c(44, 45, 52, 131, 137)){ #Indexes of selected pathways
  pathway_IFN <- data.frame(Term = enrichr_sig$Term[i], Overlap = enrichr_sig$Overlap[i], Adjusted.P.value = enrichr_sig$Adjusted.P.value[i], Genes =  enrichr_sig$Genes[i])
  
  enrichr_IFN <- rbind(enrichr_IFN, pathway_IFN)
}

rm(pathway_IFN)

```

#### Building bar plot

```{r}
enrichr_IFN$Term <- factor(enrichr_IFN$Term, levels = enrichr_IFN$Term ) #To make it the right order for ggplot

pdf("Comparison_pathways_enrichr_IFN_bar.pdf", width = 10, height = 4)
p<-ggplot(data = enrichr_IFN, aes(x = Term, y = Overlap)) + geom_bar(stat="identity", aes(fill = Adjusted.P.value)) + scale_fill_gradientn(colors = c("#668466", "#CC0000")) + coord_flip() + theme_light() + theme(axis.title.y=element_blank())

p

dev.off()

rm(p)

```




---
## Enrichr - Reactome_2016

### Import Enrichr results

https://maayanlab.cloud/Enrichr/enrich?dataset=c772337dc1494106be1825a7eae88c48

Library: Reactome_2016

"The p-value is computed from the Fisher exact test which is a proportion test that assumes a binomial distribution and independence for probability of any gene belonging to any set."

*The smaller the p-value, the higher the chance that the genes belong to that pathway, right?*

```{r}
enrichr <- read.delim("Enrichr_Reactome_2016_table.txt")

```

### Filter for only important info
```{r}
enrichr_sig <- data.frame(Term = c(), Overlap = c(), Adjusted.P.value = c(), Genes = c())

for (i in 1:nrow(enrichr)){
  if (enrichr$Adjusted.P.value[i] < 0.05) {
    sig = data.frame (Term = enrichr$Term[i], Overlap = enrichr$Overlap[i], Adjusted.P.value = enrichr$Adjusted.P.value[i], Genes = enrichr$Genes[i])
    enrichr_sig <- rbind(enrichr_sig, sig)
  }
}

#Update enrichr data.frame
enrichr <- enrichr_sig

rm(i)
rm(sig)
rm(enrichr_sig)
```

### Convert overlap into number and round to 2 decimal places

```{r}
for (i in 1:nrow(enrichr)){
  enrichr$Overlap[i] <- round(eval(parse(text = enrichr$Overlap[i])), digits = 2) #2 decimal places
}

rm(i)
#and order from top to bottom overlap
#top <- top[with(top, order(Overlap)),]

```

### Write into file

```{r}
write.xlsx(enrichr, "Enrichr_Reactome_2016_sig.xlsx", sheetName = "Enrichr_sig", col.names= TRUE, row.names = FALSE)
```

---
## Pathway network for cytoscape

### Make list of pathways and genes found from the IFN pathways

```{r}

path_genes <- data.frame(Term = c(), Gene = c())

type <- data.frame(Node = c(), Type = c())

for (i in 1:nrow(enrichr_IFN)){ #Only the 10 most significant pathways
  
  #Add gene to type table
  node <- data.frame(enrichr_IFN$Term[i], "Pathway")
  names(node) <- c("Node", "Type")
  
  type <- rbind(type, node)
  
  #Get string of genes
  genes <- strsplit(enrichr_IFN$Genes[i], ";")
  genes <- genes[[1]]
  
  for (l in 1:length(genes)){#Create an entry in the data frame for every gene
    
    #Add gene to type table
    node <- data.frame(genes[l], "Gene")
    names(node) <- c("Node", "Type")
    type <- rbind(type, node)
    
    #Add Pathway_Gene
    edge <- data.frame(enrichr_IFN$Term[i], genes[l])
    names(edge) <- c("Term", "Gene") #Data.frames should have consistent rownames
    
    path_genes <- rbind(path_genes, edge)
  }
}

#Delete temporary variables
rm(node)
rm(edge)
rm(path_genes)
```


### Write lists on files for Cytoscape

```{r}
write.xlsx(path_genes, "Cytoscape_Pathway_Genes_IFN.xlsx", sheetName = "Pathway_Genes", col.names= TRUE, row.names = FALSE)

write.xlsx(type, "Cytoscape_Type_IFN.xlsx", sheetName = "Type", col.names= TRUE, row.names = FALSE)
```


